---
title: "Cleaning Data"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

Here we'll show an example of removing reflections from an acosutic tag dataset.
In this example, each individual is marked by the unique pulse rate interval on their acoustic tag.
We'll use a HTI dataset which has been loaded and saved as an example dataset for this package.

First, we'll figure out what the pulse rate intervals of from our tags.

```{r}
## Load dataset
# Only load hydrophone 1, and all detections within first minute
dataset <- dataset_HTI
dataset.H1 <- subset(dataset, Hydrophone == 1 & seconds < 60)

## Find unique pulse rate intervals
?TagFreq.detect
pulseIntervals <- TagFreq.detect(dataset.H1$seconds,n.tags = 4,plot = T,
                     frequencies = seq(0.9,1.1,0.001), minGap = 0.0001)
```

The peaks in the histogram are not much higher than the noise floor of the plot.
This suggests that we might have to increase our frequency resolution to get better results.
You can continue fine-tuning your search frequency cector until you get a high resolution result.

```{r}
pulseFrequencies <- TagFreq.detect(dataset.H1$seconds,n.tags = 4,plot = T,
                     frequencies = seq(0.95,1.05,0.0001), minGap = 0.0001)
```

This plot looks good, although we're detecting multiple peaks in the same frequency.
You can adjust the `minGap` parameter to filter out the effects of clustered peaks.

```{r}
pulseFrequencies <- TagFreq.detect(dataset.H1$seconds, n.tags = 4, plot = T,
                     frequencies = seq(0.98,1.02,0.00001), minGap = 0.001)

pulseFrequencies
```

Now we see 4 clearly defined peaks, so we'll keep these as our unique identifiers for our individuals.

Next, we'll use the filter functions to correlate each pulse detection with the correct individual based on their tag pulse frequency.
In the plot below, red indicates detections that will be assigned to the individual.
You can use the `sensitivity` (radians) parameter to adjust your filter bandwidth.  

```{r}
require(toal)
?TagFreq.filter
tag.filter.idx <- TagFreq.filter(detections = dataset.H1$seconds,
                     frequencies = pulseFrequencies, sensitivity = 0.1, plot = T)

dataset.H1.Tag1 <- dataset.H1$seconds[tag.filter.idx[[1]]]
```

Now we've isolated the detections of individual 1 on reciever hydrphone 1, we can filter out the reflections.

```{r}

```
