---
title: "Comparison of localization methods"
author: "James Campbell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r,include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6)
```

`toal` holds a few different functions for localizing your acoustic tag data.
This vignette will demonstrate the differences between these methods.

## YAPS

YAPS (Yet Another Positioning System) was proposed by Baktoft et al. (2017) and estimates positions from tag pulses by random walk models.
This results in much more robust localization methods as compared to those which are based on traditional localization methods based soled on differences in times of arrival as compared to areference hydrophone.
We've adapted the proposed YAPS models for different situations, with a particular focus on mesocosm experiments where an acoustically tagged fish is being tracked via a three-dimentional hydrophone array in a restricted swimming area.

Some key differences between Baktoft's YAPS and the flavors of YAPS presented here is that we:

- Assume a constant speed of sound throughout the sample period.
- Assume hydrophones have errors due to guassian random movement in three-dimentions.
- Works only on 3D positioning (as opposed to only on 2D).

While the extra assumptions of the YAPS model helps with position estimate robustness, if these assumptions are not clearly met there is a risk of misleading results.
Here we'll walk through some examples testing our YAPS flavors on "simulated data".
The "simulated data" presented here consists of datasets of estimated fish positions from previous experiments.
These positions have been cleaned and we will treat these as the true positions of the simulated fishs we are working with.
Finally, we can simulate sources of error on top of our "true fish" positions.
These sources of error include:

- Guassian movement of reciever hydrophones, as assumed to be a result of wave action.
- Sample rate limitations from reciever setup

### YAPS Continual Swimming

This flavor of YAPS is based on the following assumptions:

- Displacement on x, y, and z axes have a mean of 0 and equal SDs on each axis.
- Speed of sound through water is a constant, known value.
- Hydrophone array arranged for 3D positioning (hydrophones are not positioned along a single 2D plane).
- Fixed tag pulse interval.

To test this out we'll use some previously collected data on sea bass.
We'll use a simulated hydrophone array with a shallow depth to exaggerate potential error along z axis positions.

```{r} 
require(toal)
require(ggplot2)
## Load seabass data (continual swimming)
data("Pen_SeaBass")

## Speed of sound (with simualted error)
c <- 1500
## Add some error to c
c.real <- rnorm(n = 1, mean = 1500,sd = 1)

## Get track time-or-arrival values for first fish
fish.ids <- unique(Pen_SeaBass[[1]]$Period)
xyz.true <- subset(Pen_SeaBass[[1]], Period == fish.ids[1])[1:150,c(3:5,1)]
# 5 min period

hydro.pos <- data.frame(x = c(11.15, 11.15, 3, 3),
                        y = c(11.15,3,11.15,3),
                        z = c(1.8,3,3,1.8))

toa.expected <- xyz2toa(hydrophone.positions = hydro.pos, 
               tag.positions = xyz.true,
               c = c.real)

## Simulate soruces of error
hydrophoneMovement.sd <- 0.01
pingDetection <- 0.95
ggplot(data = data.frame(ms = rnorm(n = 10000,
                                    mean = 0,
                                    sd = hydrophoneMovement.sd)*100)) + 
  geom_histogram(aes(x = ms)) + 
  ggtitle('Simulated Hydrophone movement (along direction of sound propagation)') +
  xlab('Centimeters') + theme_bw()

toa.real <- toaSimError(toa = toa.expected, reciever.fs = 12000,
                        hydrophoneMovement.sd = hydrophoneMovement.sd,
                        p.detect = pingDetection, c = c.real)

# Plot differences between expected and real toa
ggplot(data = as.data.frame(rbind(
  data.frame(Hydrophone = factor(1), ms = 1000*matrix(toa.real[,1] - toa.expected[,1], ncol = 1)),
  data.frame(Hydrophone = factor(2), ms = 1000*matrix(toa.real[,2] - toa.expected[,2], ncol = 1)),
  data.frame(Hydrophone = factor(3), ms = 1000*matrix(toa.real[,3] - toa.expected[,3], ncol = 1)),
  data.frame(Hydrophone = factor(4), ms = 1000*matrix(toa.real[,4] - toa.expected[,4], ncol = 1))))) +
  geom_histogram(aes(x = ms), bins = 40) + 
  facet_wrap(facets = ~ Hydrophone, ncol = 2) +
  ggtitle('Simulated Latency Error per Detection') +
  xlab('Milliseconds') + theme_bw()
```

Now we have our dataset, we'll apply YAPS and compare its results to that of spherical interpolation.
To make sure things don't take too long, we'll restrict YAPS to 2000 iterations.
We'll also try one with 5000 to see if that gives a noticable improvement.
More iterations should result in potentially more accurate results.
We'll also time how long each method takes.

```{r, warning=F}
## Spherical interopolation
estimates.si <- TOA.localization(toa = toa.real,
                                 hydrohpone.positions = hydro.pos,
                                 c = c)
```

We'll mute the output from this code block... as we've set YAPS to be very verbose during its run so we can check out the optimization while its running.
We've setup this function to first estimate the position using spherical interpolation.
These preliminatry results will then be used as starting values in the YAPS model to speed up the optimization.

As this model is rather complex, its important to set realistic starting parameters (via the `param` function).
Here, we've set some initial values which we know work with this dataset.
This results in the model running quite fast.
Ideally, you can initally test your model on a small subset of data to find the right parameters.

Finally, make sure you have a large enough smaple size for YAPS's assumptions to be met.
YAPS assumes that the mean displacement of this fish on the x, y, and z axis will be 0.
This will be violated if the fish swims in only one direction over your sample period.

```{r,results=F, message=F, warning=F}
## YAPS-Pelagic
params <- list(
  logSigma_bi = -7, #log transformed SD of pulse intervals
  # logSigma_dl = rep(-10,length = nrow(hydro.pos)),		# Sigma for latency error
  logSigma_dl = -10,		# Sigma for latency error
  logSigma_toa = -15, # Time of arrival SD
  logSigma_x = -3,
  logSigma_y = -3,
  logSigma_z = -3)

require(tictoc)
tic();estimates.yaps.p <- yaps.pelagic(toa = toa.real,
                                 hydrophone.pos = hydro.pos,
                                 max.iterations = 5000, params = params); toc();
```

Now we'll plot the results and compare the Spherical interpolation, which is a more conventional solution to localization.

```{r}
require(magrittr)

## Append positional data
results <- rbind(as.matrix(estimates.si[estimates.si$eq == '+',2:4]),
                 as.matrix(estimates.yaps.p$XYZ),
                 as.matrix(xyz.true[,1:3])) %>% 
  as.data.frame()

## Get results from all models in a long form data.frame for plotting
results$Model <- c(rep('SI+', dim(estimates.si[estimates.si$eq == '+',])[1]),
                   rep('YAPS.p', times = dim(estimates.yaps.p$XYZ)[1]),
                   rep('True', times = dim(xyz.true)[1]))
row.names(results) <- NULL
head(results)

lt <- 0.5
ggplot(data = results) + 
  geom_path(aes(x = x, y = y, color = Model, size = Model,
                linetype = Model, col = Model)) +
  coord_equal(xlim = range(xyz.true$PosX),
                  ylim = range(xyz.true$PosY)) +
  theme_bw() + 
  scale_size_manual(values = c('SI-' = lt,
                              'SI+' = lt,
                              'True' = lt*2,
                              'YAPS.p' = lt)) +
  scale_linetype_manual(values = c('SI-' = 1,
                              'SI+' = 1,
                              'True' = 2,
                              'YAPS.p' = 1)) +
  scale_color_manual(values = c('SI-' = 'red',
                              'SI+' = 'red',
                              'True' = 'green',
                              'YAPS.p' = 'blue')) +
  ggtitle('Horizontal Plane')

ggplot(data = results) + 
  geom_path(aes(x = y, y = z, color = Model, size = Model,
                linetype = Model, col = Model)) +
  coord_equal(xlim = range(xyz.true$PosY),
                  ylim = range(xyz.true$PosZ)) +
  theme_bw()  +
  scale_size_manual(values = c('SI-' = lt,
                              'SI+' = lt,
                              'True' = lt*2,
                              'YAPS.p' = lt)) +
  scale_linetype_manual(values = c('SI-' = 1,
                              'SI+' = 1,
                              'True' = 2,
                              'YAPS.p' = 1)) +
  scale_color_manual(values = c('SI-' = 'red',
                              'SI+' = 'red',
                              'True' = 'green',
                              'YAPS.p' = 'blue')) +
  ggtitle('Vertical Plane')
```

YAPS seems to do a good job of filling in the missing detections (when less than 4 hydrophones recieve a signal).
Additionally, YAPS seems to result in smoother paths, particularily on the z-axis which represents a worst case scenario when there was not sufficient space between the hydrophones along that axis in the experimental setup.

Finally, lets look at the distribution of our error terms, hydrophone latency.

```{r}
ggplot(data = rbind(
  data.frame(Hydrophone = 1, ms = 1000*matrix(estimates.yaps.p$dl[,1], ncol = 1)),
  data.frame(Hydrophone = 2, ms = 1000*matrix(estimates.yaps.p$dl[,1], ncol = 1)),
  data.frame(Hydrophone = 3, ms = 1000*matrix(estimates.yaps.p$dl[,1], ncol = 1)),
  data.frame(Hydrophone = 4, ms = 1000*matrix(estimates.yaps.p$dl[,1], ncol = 1)))) +
  geom_histogram(aes(x = ms), bins = 40) + 
  facet_wrap(facets = ~ factor(Hydrophone), ncol = 2) +
  ggtitle('Estimated Latency Error per Detection') +
  xlab('Milliseconds') + theme_bw()
```

The hydrophone latency is an issue here.
As seen here, 1 hydrophone typically collects all the latency error (which should be distributed evenly accross all 4 hydrophones.
Despite this, I'll keep this random effect, as in its absence, I'm guessing the locations will end up trying to compensate for this error and result in more variable resulting paths.
A next step would be to simulate this latency error using sin waves (to replicate surface waves), resulting in a more realistic form of this latency error.

## References

Baktoft, H. Gjelland, K., Økland, F., Thygesen, U.  2017. **Positioning of aquatic animals based on time-of-arrival and random walk models using YAPS (Yet Another Positioning Solver)**. *Scientific Reports*. 7:17294.
