---
title: "TOAL: Spherical Interpolation"
author: "James Campbell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### Overview

Here, we'll create a simulated hydrophone array and acoustic tag then use the functions of this package to localize the source using time-of-arrival differences.

### Create a simulated dataset

The code below will create a simulated array of 5 hydrophones and randomly place multiple sound sources within this array.

```{r, fig.width=6,fig.height=5}
require(ggplot2)
require(toal)
## Define a simulated sensor array.
sensorLocations <- data.frame(x = c(0, 0, 10, 10, 5),
                              y = c(0, 9, 0, 10, 0),
                              z = c(0, 5,5, 0, 0),
                              row.names = paste0('H',1:5))


## Create 5 random source locations within array
n <- 15 # point sources to simulate
sourceLocation <- data.frame(x = runif(n = n,min = 0, max = 10), 
                             y = runif(n = n,min = 0, max = 10), 
                             z = runif(n = n,min = 0, max = 5))

## Calculate simulated arrival delays
N <- nrow(sensorLocations) # Number of sensors
c <- 1500 # Speed of sound in water

## Function for calculating distance between points
dist <- function(a,b){
  ## Calc distance between two locations
  dif <- abs(a - b)
  return(sqrt(sum(dif^2)))
}

# Initalize empty array of TOAs
sensorLatency = matrix(ncol = N,nrow = nrow(sourceLocation))
# Calculate distances from each
for(i in 1:N){
  for(j in 1:nrow(sourceLocation)){
    sensorLatency[j,i] <- dist(sourceLocation[j,],sensorLocations[i,1:3])
  }
}
rownames(sensorLatency) <- paste0('Source ',1:nrow(sourceLocation))
colnames(sensorLatency) <- paste0('Hydrophone ',1:N)

## Convert distances to arrival latencies (from meters to seconds)
sensorLatency <- sensorLatency/c

## Plot resulting source locations and location estimates
nudge <- 0.5
ggplot(sensorLocations) + 
  geom_point(aes(x = x, y = y, size = z)) +
  geom_point(data = sourceLocation,
             aes(x = x, y = y, size = z),
             color = 'red', pch = 4) +
  geom_text(aes(x = x, y = y, label = rownames(sensorLocations)),
            nudge_x = nudge, nudge_y = nudge) +
  theme_bw() + coord_equal() +
  scale_size_continuous(name = 'Depth',trans = 'reverse') + 
  ggtitle('Simulated Sensor array and sound sources')
```

### Estimating locations

Before we run the extimate our spatial positions, we should consider that we may not know the exact speed of sound in a real dataset.
We can deal with this by using R's optimization functions to reduce the resultin error from our TOALs.
We've written a function to automate this, but the gist is that spherical interpolation is a very computationally efficient proceedure which also returns error values (spherical interpolation is essentially a closed-form solution to minimizing error already).
By letting R test different speeds of sound, we can let R choose the most likely values for us.
This is handy for addressing measurement or placement errors in your setup, and also a simple method for estimating the speed of sound from your setup without using any additional measurements (like salinity, temperature, etc).

Below, we'll type in an incorrect `c`, and the `findc` function will correct us, returning the true speed of sound.

```{r}
findc(sensorLocations = sensorLocations,detections = sensorLatency, c = 1600)
```

To localize the sound soruces, we need the time-of-arrival and hydrophone location data, seen here in `sourceLocation` and `sensorLatency`.

```{r}
sensorLocations
sensorLatency
```

The function `TOA.localization` will calculate the source locations estimates for us, along with an error parameter for each estimate.

```{r, fig.width=6,fig.height=5}
sourceEstimates <- TOA.localization(toa = sensorLatency,
                                    hydrohpone.positions = sensorLocations,
                                    c = 1500)

## Plot resulting source locations and location estimates
ggplot(sensorLocations) + 
  geom_point(aes(x = x, y = y, size = z)) +
  geom_text(aes(x = x, y = y, label = rownames(sensorLocations)),
            nudge_x = nudge, nudge_y = nudge) +
  geom_point(data = sourceLocation,
             aes(x = x, y = y, size = z),
             color = 'red', pch = 4) +
  geom_point(data = sourceEstimates,
             aes(x = x, y = y,
                 size = z),
             pch = 3, color = 'blue') +
  theme_bw() + coord_equal() +
  scale_size_continuous(name = 'Depth',trans = 'reverse') + 
  ggtitle('Simulated Sensor array')
```

In the image above, blue crosshairs represent the location estimates.
As seen, in a ideal situaiton with no error on the time-of-arrival measurements the estimates perfetly match the real source locations.


The spherical interpolation method is essentially a least squares problem for minimizing error from differences of arrival times.
Thus, for each estimate, an error parameter is calculated which can be kept in our dataset (`error` column in `sourceEstimates`).
To convert time of arrival differences to distance from source distances for each hydrophone, the spherical interpolation method requires a prior definition of the speed of sound.
Incorrectly defining this can cause large errors in your estimates.
However, with some trial and error, you can minimize the mean error to find the true speed of sound in your experimental area (or at least very close to it).
The plot below shows the error terms resulting from using incorrect sound speeds, along with a plot to visualize the result.

```{r, fig.width=6,fig.height=5}
sourceEstimates.error <- rbind(cbind(delta_c = 0,TOA.localization(toa = sensorLatency,
                                    hydrohpone.positions = sensorLocations,
                                    c = 1500)),
                               cbind(delta_c = 050,TOA.localization(toa = sensorLatency,
                                    hydrohpone.positions = sensorLocations,
                                    c = 1450)),
                               cbind(delta_c = 0100,TOA.localization(toa = sensorLatency,
                                    hydrohpone.positions = sensorLocations,
                                    c = 1400)))

## Plot resulting source locations and location estimates
ggplot(sensorLocations) + 
  geom_point(aes(x = x, y = y, size = z)) +
  geom_text(aes(x = x, y = y, label = rownames(sensorLocations)),
            nudge_x = nudge, nudge_y = nudge) +
  geom_point(data = sourceLocation,
             aes(x = x, y = y, size = z),
             color = 'red',  pch = 4) +
  geom_point(data = sourceEstimates.error,
             aes(x = x, y = y,
                 size = z, color = factor(delta_c)),
             pch = 3, stroke = 1) +
  theme_bw() + coord_equal() +
  scale_size_continuous(name = 'Depth',trans = 'reverse') + 
  ggtitle('Simulated Sensor array') +
  scale_color_discrete(name = 'c error')
```


Finally, we can test a scenrario with 4 hydrophones in the array.
This setup will cause the function `TOA.localization` to use a different formulation of the spherical interpolation method which specifically deals with 4 sensors and provides two estimates for each source location.
For this, we'll use the same simulated setup as before, but we'll remove the 5th hydrophone from the array.

```{r, fig.width=6,fig.height=5}
sensorLocations.2 <- sensorLocations[1:4,]
sensorLatency.2 <- sensorLatency[,1:4]

ggplot(sensorLocations.2) + 
  geom_point(aes(x = x, y = y, size = z)) +
  geom_point(data = sourceLocation,
             aes(x = x, y = y, size = z),
             color = 'red', pch = 4) +
  geom_text(aes(x = x, y = y, label = rownames(sensorLocations.2)),
            nudge_x = nudge, nudge_y = nudge) +
  theme_bw() + coord_equal() +
  scale_size_continuous(name = 'Depth',trans = 'reverse') + 
  ggtitle('Simulated Sensor array and sound sources')
```

A seen below, in certain setups, one type of equation will always be correct.
You can use the colum `eq` in the resulting data.frame to easily filter out the correct localizations in these situations.

```{r, fig.width=6,fig.height=5}
sensorLocations.2
sensorLatency.2

sourceEstimates.2 <- TOA.localization(toa = sensorLatency.2,
                                    hydrohpone.positions = sensorLocations.2)
ggplot(sensorLocations.2) + 
  geom_point(aes(x = x, y = y, size = z)) +
  geom_point(data = sourceLocation,
             aes(x = x, y = y, size = z),
             color = 'red', pch = 4) +
  geom_text(aes(x = x, y = y, label = rownames(sensorLocations.2)),
            nudge_x = nudge, nudge_y = nudge) +
  theme_bw() + coord_equal() +
  scale_size_continuous(name = 'Depth',trans = 'reverse') + 
  ggtitle('Simulated Sensor array and sound sources') + 
  geom_point(data = sourceEstimates.2,
             aes(x = x, y = y, size = z, color = factor(eq)),
             pch = 3, stroke = 1) + 
  scale_color_discrete(name = 'Equation')
```
