---
title: "Testing the robustness of hydrophone array detections"
author: "James Campbell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### Overview

Here, we'll try to estimate the error of our hydrophone array using simulations.
The simulation will contain error due to moving hydrophones (such as those attached to a floating platform) and the sample rate of the system.

### Create a simulated dataset

The code below will create a simulated array of 4 hydrophones and randomly place multiple sound sources within a defined area around the array.
We'll also add the error due to the hydrophone sampling rate and movement to our simulated detections, while remembering the real locations of the sources.

```{r, fig.width=8,fig.height=8,warning=FALSE}
## Define a simulated sensor array.
# Hydrophones are arranged in a rectangle with the following lateral and depth 
# distances along the edges of the array.
dist.lateral <- 11
dist.depth <- 1.4

## Define a rectangle where the sources will be randomly placed
source.lateral <- 8
source.depth <- 4

## The sampling rate of the simulated hydrophone array
# This simulate the resolution of a real hydrophone array.
fs <- 12000

## Maximum error due to hydrophone movement
# This will be used to create a random uniform distribution of hydrophopne position errors
movement.error <- 0.05
  
## Begin creating simulated dataset
require(ggplot2)
require(toal)
## Generate hydrophone positions
lat.low <- (source.lateral/2) - (dist.lateral/2)
lat.high <- (source.lateral/2) + (dist.lateral/2)
sensorLocations <- data.frame(x = c(lat.low, lat.low, lat.high, lat.high),
                              y = c(lat.low, lat.high, lat.low, lat.high),
                              z = c(0, dist.depth,dist.depth, 0),
                              row.names = paste0('H',1:4))

## Create 5 random source locations within array
n <- 400 # point sources to simulate
sourceLocation <- data.frame(x = runif(n = n,min = 0, max = source.lateral), 
                             y = runif(n = n,min = 0, max = source.lateral), 
                             z = runif(n = n,min = 0, max = source.depth))

## Calculate simulated arrival delays
N <- nrow(sensorLocations) # Number of sensors
c <- 1500 # Speed of sound in water

## Function for calculating distance between points
dist <- function(a,b){
  ## Calc distance between two locations
  dif <- abs(a - b)
  return(sqrt(sum(dif^2)))
}

# Initalize empty array
sensorLatency = matrix(ncol = N,nrow = nrow(sourceLocation))
# Calculate distances from each source
for(i in 1:N){
  for(j in 1:nrow(sourceLocation)){
    sensorLatency[j,i] <- dist(sourceLocation[j,],sensorLocations[i,1:3])
  }
}
rownames(sensorLatency) <- paste0('Source ',1:nrow(sourceLocation))
colnames(sensorLatency) <- paste0('Hydrophone ',1:N)

## Add random uniform error to each distance
errorvals.movement <- runif(min = movement.error/-2, max = movement.error/2,n = length(sensorLatency))
sensorLatency <- sensorLatency + errorvals.movement

## Convert distances to arrival latencies (from meters to seconds)
sensorLatency <- sensorLatency/c

## Apply error due to minimum sample rate
dt <- 1/fs
# dt is the sample period (i.e. minimum time between detections)
sensorLatency <- round(sensorLatency/dt,digits = 0)*dt

## Plot resulting source locations and location estimates
nudge <- 0.5
ggplot(sensorLocations) + 
  geom_point(aes(x = x, y = y, size = z)) +
  geom_point(data = sourceLocation,
             aes(x = x, y = y, size = z),
             color = 'red', pch = 4) +
  geom_text(aes(x = x, y = y, label = rownames(sensorLocations)),
            nudge_x = nudge, nudge_y = nudge) +
  theme_bw() + coord_equal() +
  scale_size_continuous(name = 'Depth',trans = 'reverse') + 
  ggtitle('Simulated Sensor array and sound sources')
```

### Calculate position estimates from simulated sources

As we're using an array of 4 hydrophones, we'll plot the results of both estimates (for different formulations of the spherical interpolation method) for each source.
To illustrate the error, we'll draw lines between each true source location and the resulting estimate from the simulated detection and error.
The interpretation is simple: long blue lines in the resulting plots indicate the magnitude and direction of estimate error.


```{r, fig.width=8,fig.height=8,warning=F}
sourceEstimates <- TOA.localization(toa = sensorLatency, 
                                    hydrohpone.positions = sensorLocations,
                                    c = c)
results <- rbind(sourceEstimates,
                 data.frame(sourceLocation,
                            id = as.numeric(rownames(sourceLocation)),
                            error = 0,
                            eq = 'real',stringsAsFactors = F))
############################################################
## Plot resulting source locations and location estimates ##
############################################################
require(gridExtra)

#### Plot sensor locations
plot.sensors.xy <- ggplot(data = results) +
  geom_point(data = sensorLocations,
             aes(x = x, y = y, size = z)) +
  geom_text(data = sensorLocations,
            aes(x = x, y = y, label = rownames(sensorLocations)),
            nudge_x = nudge, nudge_y = nudge)
plot.sensors.xz <- ggplot(data = results) +
  geom_point(data = sensorLocations,
             aes(x = x, y = z, size = y)) +
  geom_text(data = sensorLocations,
            aes(x = x, y = z, label = rownames(sensorLocations)),
            nudge_x = nudge, nudge_y = nudge)

alpha.line <- 0.5

#### Plot xy location estimates
cols <- c('#ff5555','#555555','#5555ff')
plot.1 <- plot.sensors.xy +
  geom_point(data = subset(results, subset = id %in% 'real'),
             aes(x = x, y = y, size = z), 
                 color = cols[2],
                 shape = 1) +
  theme_bw() + coord_equal() +
  geom_line(data = subset(results, subset = eq %in% c('+','real')),
            aes(x = x, y = y, group = factor(id)),
            color = cols[3], alpha = alpha.line) +
  scale_size_continuous(name = 'Depth',trans = 'reverse') + 
  ggtitle(paste0('Simulated Sensor array, M.Error: ', movement.error))


plot.2 <- plot.sensors.xy +
  geom_point(data = subset(results, subset = id %in% 'real'),
             aes(x = x, y = y, size = z), 
                 color = cols[2],
                 shape = 1) +
  theme_bw() + coord_equal() +
  geom_line(data = subset(results, subset = eq %in% c('-','real')),
            aes(x = x, y = y, group = factor(id)),
            color = cols[1], alpha = alpha.line) +
  scale_size_continuous(name = 'Depth',trans = 'reverse') + 
  ggtitle('Simulated Sensor array')

#### Plot xz location estimates
plot.3 <- plot.sensors.xz +
  geom_point(data = subset(results, subset = id %in% 'real'),
             aes(x = x, y = z, size = z), 
                 color = cols[2],
                 shape = 1) +
  theme_bw() + coord_equal() +
  geom_line(data = subset(results, subset = eq %in% c('+','real')),
            aes(x = x, y = z, group = factor(id)),
            color = cols[3], alpha = alpha.line) +
  scale_size_continuous(name = 'Y',trans = 'reverse') + 
  scale_y_reverse() +
  ggtitle('Simulated Sensor array') +
  ylab('Depth')


plot.4 <- plot.sensors.xz +
  geom_point(data = subset(results, subset = id %in% 'real'),
             aes(x = x, y = z, size = z), 
                 color = cols[2],
                 shape = 1) +
  theme_bw() + coord_equal() +
  geom_line(data = subset(results, subset = eq %in% c('-','real')),
            aes(x = x, y = z, group = factor(id)),
            color = cols[1], alpha = alpha.line) +
  scale_size_continuous(name = 'Y',trans = 'reverse') +
    scale_y_reverse() +
  ggtitle('Simulated Sensor array') +
  ylab('Depth')

  grid.arrange(plot.1, plot.2, 
               plot.3, plot.4, ncol = 2)
```
