---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

## YAPS: 2D example

```{r setup, include = FALSE}

  #Position of hydrophones. H14 malfunctioned and is not used - (x,y ) = (-9999,-9999)
  Hx = c(0,143.34,105.93,83.45,-86.3,-133.42,-107.29,15.05,36.91,-91.02,-44.82,
         61.99,113.24, -9999, 86.9,-125.92,21.67,52.63,55.33,-2.12,-41.71,-22.99,
         -78.85,-83.19,74.96,27.66,44.05,-115.85)
  Hy = c(0,-78.85,-85.87,-121.16,22.39,-55.03,31.69,-24.81,-34.47,4.96,6.59,
         -15.86,-53.76,-9999, -40.88,-33.9,-70.53,-93.06,-46.82,-24.19,-27.04,
         -60.77,-58.98,-19.39,-66.14,-5.73,-10.58,10.58)
  message('Hydrophones: ',length(Hx))

  #### Prepare TOAs
  # Must be in a matrix, where rows are hydrophones, and columns are pings
  toa <- read.table('../yaps_sciRep/toa.txt') #Time-of-arrival obtained from the hydrophones.
  toa <- t(as.matrix(toa))

  #### Subset TOAs
  #We suggest using a sub-sample for first testing (e.g. first 500 pings) as computation time is faster
  #first toa-observation to include
  nstart <- 1
  #number of pings to include - increase to 3244 for complete track. Be carefull if going below ~100 as number of detecting hydrophones is quite low in the beginning of the track (often <3 hydrophones detecting each ping); too many of these can cause numerical problems.
  n <- 1000
  nmax <- min(n + nstart - 1, ncol(toa))	#last toa-observation to include

  toa <- toa[,nstart:nmax]
  T0 <- min(toa, na.rm=TRUE)
  toa <- toa - T0


  #Increase pNA (range 0.00 - 1.00 ) to randomly increase probability of missed detections
  pNA <- 0.00
  toa[which(rbinom(length(toa),1,pNA) == 1)] <- NA

  #Subsample the hydrophone array by excluding hydros specified in noHs (range 1 - 28).
  noHs <- c(14) #Hydrophone 14 malfunctioned and is not used

  #This configuration leaves eight hydrophones to cover the entire study site
  # noHs <- c(14,28,16,10,5,8,26,27,19,18,13,3,20,9,25,24,21,22, 15,1)
  toa[noHs, ] <- NA


  
  toa[is.na(toa)] <- -9999 			#NA's are translated to -9999

  #Get data prepared for TMB
  datTmb <- list(H = matrix(c(Hx,Hy), ncol = 2), # x, y hydrophone positiosn
                 nh = length(Hx), # Number of hydrophones
                 np = ncol(toa), # Number of pings
                 toa = toa)


  ## Get list of parameters to be estimated by TMB-model
  params <- 	list(
    XY = matrix(c(mean(datTmb$H[,1]) + rnorm(ncol(datTmb$toa), sd=50),
                  mean(datTmb$H[,2]) + rnorm(ncol(datTmb$toa), sd=50)),
                ncol=2),	#positions: Use guassian random values for XY positions
    top = na.approx(
      apply(datTmb$toa,
            2,
            function(k) {median(k[k != -9999])}),
      rule=2),								#time of ping (fill in missing with interpolated values)
    ss=rnorm(ncol(datTmb$toa), 1415, 5),																			#speed of sound
    logD_xy = -2,				#diffusivity of transmitter movement (D_xy in ms)
    # logD_b = -15,				#diffusivity of burst interval (D_b in ms)
    logSigma_bi = -5,			#sigma  burst interval (sigma_bi in ms)
    logD_v = -3,				#diffusivity of speed of sound (D_v in ms)
    logSigma_toa = -8,			#sigma for Gaussian
    logScale = -3,				#scale parameter for t-distribution
    log_t_part = -3				#Mixture ratio between Gaussian and t
  )
  
    ## Get list of initial values for estimated fixed parameters
  inits <- c(-2, -5, -10, -5, -1, -2)
  names(inits) <- c('init_logD_xy', 'init_logSigma_bi', 'init_logD_v',
                    'init_logSigma_toa', 'init_logScale', 'init_log_t_part')
  inits
```

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
